Discriminating the Influence of Correlated Factors \\from Multivariate Observations: the Back-to-Back Regression

### 2020-05-19 paper/paper.tex
abstract

\begin{abstract}
Identifying causes solely from observations can be particularly challenging when
i) potential factors are difficult to manipulate independently and ii)
observations are multi-dimensional. To address this issue, we
introduce ``Back-to-Back'' regression (B2B), a linear method designed to efficiently
measure, from a set of correlated factors, those that most plausibly account for
multidimensional observations. First, we prove the consistency of B2B, its links to
other linear approaches, and show how it provides a robust, unbiased and
interpretable scalar estimate for each factor.
Second, we use a variety of simulated data to show that B2B outperforms
least-squares regression and cross-decomposition techniques (e.g. canonical
correlation analysis and partial least squares) on causal identification when
the factors and the observations are partially collinear.
Finally, we apply B2B to magneto-encephalography and fMRI of 102 subjects
recorded during a one hour long reading task to test whether our method appropriately disentangles
the respective contribution of word length and word frequency - two correlated
factors known to cause early visual and late associative cortical responses respectively. The results
show that these two factors are better disentangled with B2B than with
other standard techniques.
\end{abstract}

\begin{keyword}
  Cross-Decomposition \sep
  Feature Discovery \sep
  Magnetoencephalography \sep
  fMRI \sep
  Decoding \sep
  Encoding \sep
  Reading \sep
\end{keyword}

@@

>>>>

\begin{abstract}
Identifying causes solely from observations can be particularly challenging when
i) potential factors are difficult to manipulate independently and ii)
observations are multi-dimensional. To address this issue, we
introduce ``Back-to-Back'' regression (B2B), a linear method designed to efficiently
measure, from a set of correlated factors, those that most plausibly account for
multidimensional observations. First, we prove the consistency of B2B, its links to
other linear approaches, and show how it provides a robust, unbiased and
interpretable scalar estimate for each factor.
Second, we use a variety of simulated data to show that B2B outperforms
least-squares regression and cross-decomposition techniques (e.g. canonical
correlation analysis and partial least squares) on causal identification when
the factors and the observations are partially collinear.
Finally, we apply B2B to magneto-encephalography and functional Magnetic
Resonance Imaging of 102 subjects
recorded during a one hour long reading task to test whether our method appropriately disentangles
the respective contribution of collinear factors such as word length, word
frequency respectively known to cause early visual and late associative cortical
responses. The results
show that these factors are better disentangled with B2B than with
other standard techniques.
\end{abstract}

\begin{keyword}
  Cross-Decomposition \sep
  Feature Discovery \sep
  MEG \sep
  fMRI \sep
  Decoding \sep
  Encoding \sep
  Reading \sep
\end{keyword}

@@

% Here, we focus on the restricted issue of disentangling the
% influence of linearly correlated predictors ($X$) onto noisy multivariate
% observations ($Y$). The present approach thus differs from other causal
% discovery algorithms based on temporal-delays and/or nonlinear interactions in
% systems where the directionality of causation (from X to Y or vice versa) is
% unknown (e.g. \citep{peters2017elements, granger1969investigating,
% janzing2013quantifying, scholkopf2016modeling}.
